---
title: "Article Analysis Notebook"
author: "Maria C. Codlin"
output:
  html_notebook: default
  html_document: default
  pdf_document: default
---
###Table of contents?

###Introduction to this document, the contents, and the article.

First I modified the files slightly to import with curl, as it refused to bring in the header names with certain symbols in (it worked fine importing from local repo). 


To start with a blank slate
```{r clear workspace}
rm(list=ls())
```


#Step 1. Load packages and data
Before I run any parts of the analysis, I load all the packages I use in this file
```{r load libraries}
require(dplyr)
require(ggplot2)
require(stringr)
require(plotrix) #to calculate standard error
require(qwraps2)
require(tidyr)
require(stargazer)
require(knitr)
```
First in excel, I deleted two rows of non-table data, i.e. title, and a note, and saved the file (S2_Table) as a CSV in the project directory. Then I read it into R and converted it to a table using dplyr. I trimed white space at beginnings and ends of things where there shouldn't be any, and kept strings as is to avoid issues later down the line. 
```{r readfile arch}
library(curl)
f <- curl("https://raw.githubusercontent.com/mazattack/Article-analysis/master/S2_Table.csv")
iso_total <- read.csv(f, header = TRUE, stringsAsFactors = FALSE, strip.white=TRUE)
```
```{r}
iso_total<-tbl_df(iso_total)
head(iso_total)
str(iso_total)
```

```{r readfile modern}

f <- curl("https://raw.githubusercontent.com/mazattack/Article-analysis/master/S1_Table.csv")
modern<-read.csv(f, header=TRUE, sep = ",", skip=0, strip.white=TRUE, stringsAsFactors = FALSE) #here I used skip=3 to skip the first 3 rows before the data started

modern<-tbl_df(modern)
head(modern)
str(modern)
```

To make life easier, I changed the names of two key variables. First I tried 18O and 13C as the variable names, but it threw me an error, I think by starting with a number. 
```{r rename}
names(iso_total)[12]<-"C13ap" #I used names(iso_total) to get the position of these for renaming
names(iso_total)[13]<-"O18ap"
names(iso_total)[1]<-"lab_n"
names(modern)[10]<-"modC13"
names(modern)[11]<-"modO18"
```


#Step 2: Remove of samples potentially affected by diagenisis and summarize data to check against means presented in the article 

In this article, they tested a random sample (n=101) of bone for potential post-depositional alteration to the mineral structure.Using FTIR, they tested the carbonate to phospate ratios (C/P) and infrared splitting factor (IR-SF) using FTIR and removed samples which fell outside of the expected ranges of 0.1 and 0.5 for C/P and 2.0 and 4.0 for IR-SF. The samples that they removed are marked in grey in the supplimentary data. 

Here, I used the filter function to select only those rows where C/P and IR-SF fell within the expected ranges, and saved this as a new dataset "iso". I also needed to make sure the rows with NA values were still included, since only a random sample was analysed. I did this using the | notation which functions like "OR". Since the FTIR analysis returned both C/P and IR-SF values, I only need to include one is.na function here. I also excluded an outlier with an extremely low delta O18 that they excluded by converting this number to NA, as the samples delta C13 value was still incorperated in the analysis.
```{r filter out diagenisis}
#to filter out only Lepus and Sylvilagus genus, I first needed to define the column as a factor
iso_total$Genus<-as.factor(iso_total$Genus)
iso<-filter(iso_total, (Genus=="Lepus" | Genus=="Sylvilagus") & between (C.P, 0.1, 0.5)&between (IR.SF, 2.0, 4.0) | (is.na(IR.SF))) #brackets around the Genus filter were important here, otherwise the | worked on everything that came after
iso$O18ap[iso$O18ap==11.9]<-NA #couldn't figure out how to get this to work with dplyr or chaining
iso            
```
```{r}
iso<-filter(iso, !(is.na(iso$Genus))) #removes empty rows
```

**Descriptive statistics from the article**
Archaological specimens: δ13Capatite = -7.8 ± 2.4‰(N = 114, 1 S.D.) and δ18Oapatite = 26.0‰(N = 113, 1 S.D.). 
Modern reference specimens: δ13Capatite = -12.4 ± 1.8‰(N = 13, 1 S.D.) and δ18Oapatite = 26.1‰± 1.4 (N = 13, 1 S.D.)."(Somerville et al. 2016:10). *I also note that they should have used "n" here*

Note: the first time I tried this it was not a match and so I realised I needed to keep the na values in the filter above. The second time it was also not correct so I added two lines to count the number of values in the two columns of interest to see if the sample sizes match. I noted the single value from delta O18 to remove but I still have one extra row in there than their sample number. I checked on the excel, and there is no evidence of this extra row that they excluded. I took a guess that this may have been a species listed in the spreadsheet as Lepus? and removal of this specimen now produces averages that match the data presented in the article. 

```{r sum_table}
my_sum <- 
  list("Archaeological samples" = 
        round (c("δC13 n" = sum(!is.na(iso$C13ap)),
              "δC13 mean" =  mean(iso$C13ap, na.rm=TRUE),
              "δC13 sd" =  sd(iso$C13ap, na.rm=TRUE),
              "δO18 n" = sum(!is.na(iso$O18ap)),
              "δO18 mean" =  mean(iso$O18ap, na.rm=TRUE),
              "δO18 sd" =  sd(iso$O18ap, na.rm=TRUE)),2),
       "Modern samples" = 
         round(c("δC13 n" = sum(!is.na(modern$modC13)),
              "δC13 mean" =  mean(modern$modC13, na.rm=TRUE),
              "δC13 sd" =  sd(modern$modC13, na.rm=TRUE),
              "δO18 n" = sum(!is.na(modern$modO18)),
              "δO18 mean" =  mean(modern$modO18, na.rm=TRUE),
              "δO18 sd" =  sd(modern$modO18, na.rm=TRUE)),2)
       )
    
my_sum
```

#Step 3 "Independent samples t-tests were conducted to compare mean stable isotope ratios of the two genera (Lepus and Sylvilagus) across the archaeological bone sample as a whole. No significant difference between the genera were found in δ13Capatite values (t[110] = -0.806, p = 0.422) or δ18Oapatite values (t[110] = 0.428, p = 0.669)."

First, I will check that the data fits a normal distribution by comparing it to the qnorm and qqline plots as well as histogram.

**Carbon Apatite Results**
*Histogram*
```{r}
a_hist<-ggplot(iso, aes(C13ap))
a_hist+geom_histogram(bins= 15)+ facet_wrap(~Genus, scales="free")
```
Sylvilagus looks approximately normal while Lepus it is hard to tell. Next I compare using qqplots.
*QQplots*
```{r qqplots carbon}
LC<-iso$C13ap[iso$Genus=="Lepus"]
SC<-iso$C13ap[iso$Genus=="Sylvilagus"]
qqnorm(LC,xlab="Theoretical",main="Lepus")
  qqline(LC)
  
  qqnorm(SC,xlab="Theoretical",main="Sylvilagus")
  qqline(SC)
  
  shapiro.test(LC)
  shapiro.test(SC)
```
Again, Lepus fits this "OK" while Sylvilagus looks pretty good. Both samples pass the Shapiro-Wilk test for normality. 
*Test of equal variance*
```{r}
var.test(LC, SC)
```

This suggests that the variance of the two samples is comparable and thus a two sample T test can be used.First I report the tests for the oxygen data and then provide both t-tests together with a discussion.


**Oxygen Apatite Results**
*Histogram*
```{r}
a_hist<-ggplot(iso, aes(O18ap))
a_hist+geom_histogram(bins=15)+ facet_wrap(~Genus, scales="free")
```
Both look like they could be approximately normal
*QQplots*
```{r qqplots Oxygen}
LO<-iso$O18ap[iso$Genus=="Lepus"]
SO<-iso$O18ap[iso$Genus=="Sylvilagus"]
qqnorm(LO,xlab="Theoretical",main="Lepus")
  qqline(LO)
  
  qqnorm(SO,xlab="Theoretical",main="Sylvilagus")
  qqline(SO)
```
Here, Lepus fits the normal distrbution well, while Sylvilagus is divergent. Looking at the histgram again, it could poentitally be bimodal. 
*Test of equal variance*
```{r}
var.test(LO, SO)
```
*Carbon apatite two sample T-test*
```{r}
t.test(LC, SC, var.equal =TRUE, alternative = "two.sided")
```




*Oxygen apatite two sample T-test*
```{r}
t.test(SO, LO, var.equal =TRUE, alternative = "two.sided")
```
*Summary:*
Carbon apatite t-sample t-test comparing Lepus and Sylvilagus: t = 0.5185, df = 112, p-value = 0.6051
Oxygen apatite t-sample t-test comparing Lepus and Sylvilagus: t = -0.83133, df = 111, p-value = 0.4076
*Values reported in Somerville et al. (2016:10)*
δ13Capatite values (t[110] = -0.806, p = 0.422) 
δ18Oapatite values (t[110] = 0.428, p = 0.669). 

Both of these tests returned different test statistics, degrees of freedom and p-values to those reported by Somerville et al. (2016). I tried multiple variations of the t-test in R, from Welsh t-test and one sided t-tests to one sided Welsh t-tests and none of the tests produced a similar result to those reported by Somerville. In fact, the two sample t-test for equal varience that I ran for both carbon and oxygen data matches most closely with the result reported for the opposite isotopic data. i.e. my oxygen looks most similar to their carbon. I have a number of hypothesis for why these tests may vary: the SPSS test is vastly different to the R t-tests, they are excluding samples from the test and not reporting it, as indicated by the difference in df between our tests, or there was an error in their reporting of the data. That said, the outcome of all tests is the same, there is no significant difference in the means of the samples. I later realised that the data they provide is slightly differerent from the results presented and this is likely the reason for the differences observed in this test. This is most clearly observed in the table below and comparison to their table 2. 

#Step 4 
"We also evaluated stable isotope ratios between the genera and between individual site locations using a two-way ANOVA (factorial ANOVA with two factors), testing for the main effects of genus and context, and for the genuscontext interaction."


*Summary table (Compare Table 2)*

```{r create genus column in modern}
Genus<- as.factor(modern$Species)
modern<-cbind(modern, Genus)
levels(modern$Genus)[levels(modern$Genus)=="Lepus californicus texianus"]<-"Lepus"
levels(modern$Genus)[levels(modern$Genus)=="Sylvilagus floridanus"]<-"Sylvilagus"
levels(modern$Genus)[levels(modern$Genus)=="Sylvilagus floridanus orizabae"]<-"Sylvilagus"
levels(modern$Genus)    
```


```{r merge modern and iso datasets}
new<-select(iso, lab_n, Site, Phase, Genus, C13ap, O18ap)
new2<-select(modern, lab_n=Lab_num, Site=State, Phase=Collection.date..yrs.AD., Genus, C13ap=modC13, O18ap=modO18)%>%mutate(Site="modern")
iso2<-rbind(filter(new,(!(is.na(C13ap)))), new2)
iso2
```
```{r constructing the summary table}
iso2$Genus <-as.factor(iso2$Genus)
iso2_g <- group_by(iso2, Site, Genus)
sum_tot<- summarize(iso2_g, length(C13ap), mean(C13ap), sd(C13ap), length(O18ap), mean(O18ap), sd(O18ap))

tot_tot<-iso2%>%
  group_by(Genus)%>%
  summarize(length(C13ap), mean(C13ap), sd(C13ap), length(O18ap), mean(O18ap), sd(O18ap))
tot_tot<-mutate(tot_tot, Site="Total")
tot<-rbind.data.frame(sum_tot, tot_tot)
tot
```
Note the differences across these summary statistics compared to the table provided in the article (Table 2). This reflects a difference in the data from the article compared to what they provided. I checked this in excel to confirm. 

```{r}
carbon<-lm(iso$C13ap~iso$Genus*iso$Site)
anova(carbon) #not sure why but moon pyramid is missing from all these ANOVA
```
```{r}
oxygen<-lm(iso$O18ap~iso$Genus*iso$Site)
summary(oxygen)
```
My two way Anova summaries match the conclusions they derived in the article. That Genus had no significant effect on either δ13C or δO18 apatite ratios. However there is a significant effect of Site location on these variables. 

**One way ANOVA tests**
```{r}
carbon<-lm(iso$C13ap~iso$Site)
summary(carbon)
```

```{r}
oxygen<-lm(iso$O18ap~iso$Site)
summary(oxygen)
```

```{r}

```



, the main effect for genus was not significant (F[1,102] =
3.361), p = 0.070), but the main effect of site context was significant (F[4,102] = 5.365,
p = 0.001). Importantly, the interaction between genus and site was not significant for δ13Capatite
(F[4,102] = 0.590, p<0.670) or for δ18Oapatite (F[4,102] = 0.834, p<0.507). However, Lepus and
Sylvilagus values are notably different from each other at the Cuevas site, but the small sample
size from this context makes any interpretations regarding this observation speculative. Subsequent
contextual comparisons thus combine Sylvilagus and Lepus specimens

*For δ13Capatite, the main effect for genus was not significant (F[1,102] =
3.361), p = 0.070), but the main effect of site context was significant (F[4,102] = 5.365,
p = 0.001). Importantly, the interaction between genus and site was not significant for δ13Capatite
(F[4,102] = 0.590, p<0.670) or for δ18Oapatite (F[4,102] = 0.834, p<0.507). 
However, Lepus and
Sylvilagus values are notably different from each other at the Cuevas site, but the small sample
size from this context makes any interpretations regarding this observation speculative. Subsequent
contextual comparisons thus combine Sylvilagus and Lepus specimens*